MS_Intro_Short

### 2023-01-28 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
moved items from disucussion to introduction

In this paper, we introduce a framework for developing a modular youth mental health MOSHEM. Specifically we:

@@

(ii) provide a worked example of how we used the framework to develop and apply MOSHEM modules for utility mapping.

>>>>

In this paper, we introduce a framework for developing the ready4 MOSHEM. Specifically we:

@@

(ii) provide a worked example of how we used the framework to develop and apply ready4 modules for utility mapping.

### 2023-01-28 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
moved items from disucussion to introduction

In this article we introduced a framework that we developed to help us implement a MOSHEM in youth mental health that is accountable, reusable and updatable. We outlined the standards for MOSHEMs specified by that framework and described the modelling toolkit we for applying those standards to model development and use. We also provided an overview of an initial set of MOSHEM modules we developed with the framework an applied to implement a utility mapping study and reviewed the modules, datasets and analyses from that study against framework standards.

The work we have described raises a number of issues relating to the future development of a youth mental health MOSHEM, for health economists interested in undertaking other MOSHEMS and for health research funders.

## Implications for the development of a youth mental helath MOSHEM
Mental disorders impose high health, social and economic burdens worldwide [@RN8;@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equipped for this challenge[@RN22;@RN23]. There is therefore an opportunity for health economists to undertake work in mental health of significant potential public benefit. To date, economic studies in mental health have predominantly examined the affordability and value for money of individual programs [@RN34]. This economic evaluation work is an essential prerequisite for developing economically sustainable mental health reform plans, but needs to be accompanied by and integrated with a broader program of economic research.

We hope that the MOSHEM we are developing will facilitate a synthesis of multiple strands of economic research in youth mental health.  Initially, we aim to undertake modelling research in four of the twelve domains of health economics from the classification system developed by Wagstaff and Culyer [@wagstaff2012four] :

- health and its value (utility mapping models);

- determinants of health and ill-health (models for creating rich and representative synthetic household populations from survey data);

- demand for health and health care (spatial epidemiology and help-seeking choice models); and

- supply of health services (a model of primary mental health care services).

We then want to be able to flexibly combine these models to answer questions in two other domains:

- efficiency and equity (optimising targeting of care and assessing the distributional impacts of candidate policies); and

- economic evaluation (assessing cost-utility of competing policy options).

When planning how to undertake and synthesise this program of work, we became concerned with issues relating to validity and granularity. We know that mistakes are easy to make in individual health economic modelling projects [@Radeva2020], an issue that is likely to be compounded when combining multiple projects. Similarly, the likelihood of certain types of error can increase if a model is too simple or too complex [@Saltelli2019], so the level of granularity appropriate for one model application may be ill-suited to another. These concerns led us to seeking an approach that was both open source (to provide the transparency that makes it easier for errors to be uncovered) and modular (to facilitate interchangeability between representations of different levels of granularity). However, because MOSHEMs are not common in health economics, we were unable to identify one up to date and consolidated source of guidance about how to undertake this work. We developed the framework's standards to fill this gap, choosing the framing of standards rather than guidelines because standardisation is an essential enabler of the interoperable and collaborative features of MOSHEMS. The framework's standards make explicit our goal to implement an accountable, reusable and updatable MOSHEM and provide a mechanism to assess how well this goal is being met. We have partially automated many of the steps for meeting those standards through the tools we developed for the framework's modelling toolkit and have been able to demonstrate the practical application of this toolkit through its use in implementing a utility mapping study [@Hamilton2021.07.07.21260129]. This toolkit is expected to be a source of significant efficiency savings as we implement other projects in the current phase of development of our youth mental health MOSHEM.

>>>>

In this article we introduced a framework that we developed to help us implement ready4 - a MOSHEM in youth mental health. We outlined framework standards for an accountable, reusable and updatable MOSHEM and described the modelling toolkit we for applying those standards to the development and use of ready4. We also provided an overview of an initial set of ready4 modules that we developed with the framework and which we applied to implement a utility mapping study in a clinical mental health sample of young people. We reviewed the modules, datasets and analyses generated by that study against framework standards.

The strengths. limitations and issues raised by this work can be discussed in terms of relevance to future development of the ready4 MOSHEM, to health economists interested in undertaking other MOSHEMS and to funders of health economic research.

## Implications for the development of the ready4 MOSHEM
Open source approaches have been recommended to help develop the mental health modelling field [@RN73] but only one mental health related model (in Alcohol Use Disorder [@Basu2018]) is currently indexed [RECHECK] in the Open Source Models Clearinghouse [@OSMC_20xx;@Emerson2019]. We  are aware of just one other open source mental health model - a reference model in Major Depressive Disorder - that is currently in development [@IVIMDD2022].

One of the main novel contributions of the framework we have developed is that it declares an explicit set of standards for the implementation and use of a MOSHEM. Standardisation is an essential enabler of workflow automation, module interoperability, and transferability of methods<!-- and efficient collaboration  -->. The framework's modelling toolkit has enabled us to partially automate workflows for authoring, documenting and disseminating ready4 modules, datasets and analyses. The modelling toolkit's practical utility has been demonstrated through its application in creating all ready4 components used in a utility mapping study [@Hamilton2021.07.07.21260129] and the automation it facilitates makes the development and curation of a large and growing number of ready4 digital assets more tractable. The interoperability of the ready4 modules we have developed today is demonstrated in multiple example literate programs on the ready4 documentation website [@rfwn2022] and in the program used to implement the utility mapping analysis [@hamilton_matthew_2022_6129906].

The framework's standards also provide a mechanism to assess the extent to which ready4 meets explicit accountability, reusability and updatability goals.

### 2023-01-28 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
moved items from disucussion to introduction

We developed a toolkit to help us develop and use a MOSHEM that meets the standards listed in Table \@ref(tab:timelygls). The toolkit is comprised of online repositories and software.

@@

For hosting documentation websites for the code libraries we create, we enabled GitHub Pages in each library's repository. We also developed a consolidated project documentation website (https://www.ready4-dev.com), using the Hugo framework [@hugo_2023] and Docsy theme [@docsy_2023]. To host that website we established an account with Netlify [@netlify_2023] and linked that service to our GitHub organisation so that the website would automatically update whenever the source code in its GitHub repository was edited.

@@

The foundation for our MOSHEM is provided by the ready4 library [@ready42022] that defines a template module (using R's S4 type class) from which all model module data structures will inherit features and a novel syntax for implementing the algorithms attached to model modules. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs and for partially automating updates to the project website.

@@

The ready4use [@ready4use2022] library contains modules for ingesting model datasets from online repositories (hosted on a Dataverse installation or on GitHub), labelling model datasets and sharing model datasets via online repositories. The ready4show library [@ready4show2022] contains tools to help author analyses programs that are either self-documenting or which trigger the creation of a scientific summary.

>>>>

We developed a toolkit to help us develop and use ready4 modules, datasets and analyses that meet the standards listed in Table \@ref(tab:timelygls). The toolkit is comprised of online repositories and software.

@@

For hosting documentation websites for the code libraries we create, we enabled GitHub Pages in each library's repository. We also developed a consolidated project documentation website [@rfwn2022], using the Hugo framework [@hugo_2023] and Docsy theme [@docsy_2023]. To host that website we established an account with Netlify [@netlify_2023] and linked that service to our GitHub organisation so that the website would automatically update whenever the source code in its GitHub repository was edited.

@@

The foundation for ready4 is provided by a library called ready4 [@ready42022] that defines a template module (using R's S4 type class) from which all model module data structures will inherit features and a novel syntax for implementing the algorithms attached to model modules. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs and for partially automating updates to the project website.

@@

The ready4use library  [@ready4use2022] contains modules for ingesting model datasets from online repositories (hosted on a Dataverse installation or on GitHub), labelling model datasets and sharing model datasets via online repositories. The ready4show library [@ready4show2022] contains tools to help author analyses programs that are either self-documenting or which trigger the creation of a scientific summary.

### 2023-01-28 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
moved items from disucussion to introduction

We plan to develop a modular OSHEM (MOSHEM) of youth mental health that is accountable, reusable and updatable. Open source approaches have been recommended to help develop the mental health modelling field [@RN73] but only one mental health related model (in Alcohol Use Disorder [@Basu2018]) is currently indexed [RECHECK] in the Open Source Models Clearinghouse [@OSMC_20xx;@Emerson2019]. We  are aware of just one other open source mental health model - a reference model in Major Depressive Disorder - that is currently in development [@IVIMDD2022].

>>>>

Our interest in modular and open source source approaches developed when we began seeking am appropriate framework for undertaking and validly synthesising diverse types of economic research in mental health. Mental disorders impose high health, social and economic burdens worldwide [@RN8;@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equipped for this challenge [@RN22;@RN23]. A substantial economic literature already exists to assess the affordability and value for money of mental health interventions [@RN34]. This economic evaluation work is an essential prerequisite for improving allocative efficiency in mental health, but could be of greater value to systems planners if integrated with a broader program of economic research.

We are developing ready4, a modular OSHEM (MOSHEM) on the mental health of young people aged 12 to 25 that we are initially applying to four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]:
<!-- There is therefore an opportunity for health economists to undertake mental health research of significant potential public benefit. -->

- *health and its value* (our projects: utility mapping models);

- *determinants of health and ill-health* (our projects: models for creating rich and representative synthetic household populations from survey data);

- *demand for health and health care* (our projects: spatial epidemiology and help-seeking choice models); and

- *supply of health services* (our projects: a model of primary mental health care services).

Once these projects are completed, our aim is to flexibly combine these models to answer questions in two additional Wagstaff and Culyer domains:

- *efficiency and equity* (our goal: assess the distributional impacts and identify the optimal targeting of care provision); and

- *economic evaluation* (our goal: assess the cost-utility of competing policy options).

Standardisation is an essential enabler of workflow automation, module interoperability, and transferability of methods<!-- and efficient collaboration  -->.

<!-- When planning how to undertake and synthesise this program of work, we became concerned with issues relating to accountability, reusability and updatability. If mistakes are common in individual health economic modelling projects [@Radeva2020], then combining multiple projects may be particularly error prone. The risk of errors may also depend on whether a model is too simple or too complex [@Saltelli2019], so the level of granularity appropriate for one model application may be ill-suited to another. Undertaking a multiannual program of interdependent work raises the risk that models developed in the first phase of a project may need to be updated by the time the project is approaching conclusion. -->

<!-- These concerns led us to seek the MOSHEM approach that provides the transparency to make it easier for errors to be uncovered, the adaptability to fix these errors and the flexibility to facilitate interchangeability between representations of different levels of granularity.  -->
